---
title: "Selecting Columns to Display in AE Listings"
authors: "Yujie Zhao"
output:
  rmarkdown::html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
vignette: |
  %\VignetteIndexEntry{Selecting Columns to Display in AE Listings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(forestly)
library(metalite)
```

The interactive AE forest plots provide drill-down capability to subject-level AE listings. In this vignette, we demonstrate how to customize the columns of the AE listing.



# Step 1: build your metadata

Building interactive AE forest plots starts with constructing the metadata. The detailed procedure for building metadata is covered in the vignette [Generate Interactive AE Forest Plots with Drill Down to AE Listing](https://merck.github.io/forestly/articles/forestly.html). Therefore, in this vignette, we will skip those details and directly use the metadata created there. 

```{r}
adsl <- forestly_adsl
adae <- forestly_adae

adsl$TRTA <- factor(forestly_adsl$TRT01A, 
                    levels = c("Xanomeline Low Dose", "Placebo"), 
                    labels = c("Low Dose", "Placebo"))
adae$TRTA <- factor(forestly_adae$TRTA, 
                    levels = c("Xanomeline Low Dose", "Placebo"), 
                    labels = c("Low Dose", "Placebo"))

meta <- meta_adam(population = adsl, observation = adae) |>
  define_plan(plan = plan(analysis = "ae_forestly", 
                          population = "apat",
                          observation = "apat", 
                          parameter = "any;drug-related")) |>
  define_analysis(name = "ae_forestly", label = "Interactive Forest Plot") |>
  define_population(name = "apat", group = "TRTA", id = "USUBJID",
                    subset = SAFFL == "Y", label = "All Patient as Treated") |>
  define_observation(name = "apat", group = "TRTA", 
                     subset = SAFFL == "Y", label = "All Patient as Treated") |>
  define_parameter(name = "any", 
                   subset = NULL,
                   label = "Any AEs", 
                   var = "AEDECOD", soc = "AEBODSYS") |>
  define_parameter(name = "drug-related", 
                   subset = toupper(AREL) == "RELATED",
                   label = "Drug-related AEs", 
                   var = "AEDECOD", soc = "AEBODSYS") |>  
  meta_build()
```

# Step 2: customize the AE listing columns

Users can specify the AE listing columns using the `ae_listing_display = ...` argument in the `prepare_ae_forestly()` function. The columns displayed in the AE listing correspond to the variables provided in `ae_listing_display = ....`.

In the following example, we select the unique subject ID, site number, gender, race, and age, so the listing will display only these five variables.

```{r}
meta |>
  prepare_ae_forestly(ae_listing_display = c("USUBJID", "SITEID", "SEX", "RACE", "AGE")) |>
  format_ae_forestly() |>
  ae_forestly()
```