---
title: "Interactive Forest Plot"
output:
  rmarkdown::html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
vignette: |
  %\VignetteIndexEntry{Interactive Forest Plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  echo = TRUE
)
```

```{r, message=FALSE}
library(metalite)
library(forestly)
```

# Overview

The forestly package creates interactive forest plots for clinical trial analysis & reporting.

- Safety analysis
  - Specific adverse events analysis
- Efficacy analysis (future work)
  - Subgroup analysis

We assume ADaM datasets are ready for analysis and
leverage [metalite](https://merck.github.io/metalite/) data structure to define
inputs and outputs.

```{r}
meta_forestly(
  dataset_adsl = forestly_adsl,
  dataset_adae = forestly_adae,
  population_term = "apat",
  observation_term = "wk12"
) |>
  prepare_ae_forestly(parameter = "any;rel;ser") |>
  format_ae_forestly() |>
  ae_forestly()
```

# Interactive features

The interactive features of the example include:

- Select different AE criteria.
- Filter by incidence of AE in one more more group.
- Revealed information by hovering the mouse over a point.
- Search bars to find subjects with selected adverse events (AE).
- Sort value by click the column header.
- Drill-down listing by clicking $\blacktriangleright$.

# Workflow

The general workflow is:

1. `meta_forestly()` constructs input metadata for treatment analysis from ADaM datasets.
1. `prepare_ae_forestly()` prepares datasets for interactive forest plot.
1. `format_ae_forestly()` formats output layout.
1. `ae_forestly()` generates an interactive forest plot.

Here `meta_forestly()` is a wrapper function of the metalite package.
The function is for users to simplify the process to define input parameters required
to generate the interactive forest plot.

`meta_forestly` function require `dataset_adsl`, `dataset_adae`, `observation_term` and `population_term` arguments. `dataset_adsl`, `dataset_adae` are the input ADaM dataset for population and observation. In this example, we will use "apat" for `population_term` and "wk12" for `observation_term`. For subset creteria for both terms, we are using `SAFFL == "Y"` by default in `meta_forestly()`.

```{r}
meta_forestly(
  dataset_adsl = forestly_adsl,
  dataset_adae = forestly_adae,
  population_term = "apat",
  observation_term = "wk12"
)
```

If user wants to customize these definitions of terms or define new terms, more tutorials can be found at this link.

For paramter level metadata, `parameter_term` is selected from predefined terms listed below. By default, we choose three parameter terms `rel`, `any` and `ser` and use ";" to concatenate them to allow the three types of AE to be displayed in one interactive forest plot.

| Term         | Subset Criteria | Label |
| :------------- |:------------- | :-----|
| `any`      | `NULL` | any adverse events |
| `rel`      | `toupper(AREL) == "RELATED"` | drug-related adverse events |
| `g34` | `ATOXGRN %in% c(3, 4)`| grade 3-4 adverse events |
| `g340rel` | `ATOXGRN %in% c(3, 4) & AREL == "Y"`| drug-related grade 3-4 adverse events |
| `g35` | `ATOXGRN %in% c(3, 4, 5)`| grade 3-5 adverse events |
| `g350rel` | `ATOXGRN %in% c(3, 4, 5) & AREL == "Y"`| drug-related grade 3-5 adverse events |
| `nonser` | `AESER != "Y" | is.na(AESER)`| non-serious adverse events |
| `ser` | `AESER == "Y"`| serious adverse events |
| `ser0rel` | `AESER == "Y" & AREL == "Y"`| serious drug-related adverse events |
| `mod` | `toupper(AEACN) %in% c("DOSE REDUCED", "DRUG INTERRUPTED", "DRUG WITHDRAWN")`| adverse events result in dose modification |
| `dth` | `AESDTH == "Y"`| adverse events result in death |
| `dtc0rel` | `AESDTH == "Y" & AREL == "Y"`| drug-related adverse events result in death |
| `disc` | `toupper(AEACN) == "DRUG WITHDRAWN"`| adverse events resulting in discontinuation |
| `disc0drel` | `toupper(AEACN) == "DRUG WITHDRAWN" & AREL == "Y"`| drug-related adverse events resulting in discontinuation |
| `disc0ser` | `toupper(AEACN) == "DRUG WITHDRAWN" & AESER == "Y"`| serious adverse events resulting in discontinuation |
| `disc0ser0rel` | `toupper(AEACN) == "DRUG WITHDRAWN" & AESER == "Y" & AREL == "Y"`| serious drug-related adverse events resulting in discontinuations |
| `inj` | `toupper(AECAT) == "I"`| injection-site |
| `noninj` | `toupper(AECAT) != "I" | is.na(AECAT)` | non-injection-site|
| `inj0rel` | `toupper(AECAT) == "I" & AREL == "Y"` | drug-related injection-site |
| `noninj0rel` | `(toupper(AECAT) == "I" | is.na(AECAT)) & AREL == "Y"` | drug-related non-injection-site |

In the following example, we define `ITT` with subset criteria `ITTFL == "Y"`for population and `TEAE` with subset criteria `TRTEMFL == "Y"` for observation. For parameter, we choose `ser` and `nonser` from pre-defined parameter list.

```{r}
meta <- meta_forestly(
  dataset_adsl = forestly_adsl,
  dataset_adae = forestly_adae,
  population_term = "ITT",
  population_subset = ITTFL == "Y",
  observation_term = "TEAE",
  observation_subset = TRTEMFL == "Y",
  parameter_term = "any;ser;nonser;rel"
)
meta
```

After the metadata is defined, we can create an interactive forest plot as below.

In this section, we calculate statistics and prepare plotting dataset using `prepare_ae_forestly()`. Argument `parameter` is required since we can select parameter from the one we defined in metadata.

Then we could use `format_ae_forestly()` to format the output and `ae_forestly()` to create the interactive forest plot. Default value is used for all arguments in these two functions.

In the first example, we select `any`, `ser` and `nonser`.

```{r}
meta |>
  prepare_ae_forestly(parameter = "any;ser;nonser") |>
  format_ae_forestly() |>
  ae_forestly()
```

In this example we select `any`, `rel` and `ser`.

```{r}
meta |>
  prepare_ae_forestly(parameter = "any;rel;ser") |>
  format_ae_forestly() |>
  ae_forestly()
```

# Define metadata

```{r}
library(metalite)
```

In order to provide more flexibility for different analysis,
The forestly package also allows other way to construct metadata.
We only show a simplified example for the metalite package.
More details can be found in the [**metalite**](https://merck.github.io/metalite/articles/index.html) package website.

Then we start to construct a metadata using `metalite`. The output metadata contains two observation `wk12` and `period1` as illustration purpose. Also, for parameter, the subset criteria is `AEREL %in% c("POSSIBLE", "PROBABLE")` instead of `toupper(AREL) == "RELATED"` in order to fit the observation dataset.

```{r}
meta <- meta_adam(
  population = forestly_adsl,
  observation = forestly_adae
) |>
  define_plan(plan = plan(
    analysis = "ae_forestly",
    population = "apat",
    observation = c("wk12", "period1"),
    parameter = "any;rel;ser"
  )) |>
  define_population(
    name = "apat",
    group = "TRT01A",
    subset = quote(SAFFL == "Y")
  ) |>
  define_observation(
    name = "wk12",
    group = "TRTA",
    subset = quote(SAFFL == "Y"),
    label = "Weeks 0 to 12"
  ) |>
  define_observation(
    name = "period1",
    group = "TRTA",
    var = "AEDECOD",
    subset = quote(TRTEMFL == "Y" & SAFFL == "Y"),
    label = "Period 1"
  ) |>
  define_parameter(
    name = "rel",
    subset = quote(AEREL %in% c("POSSIBLE", "PROBABLE"))
  ) |>
  define_analysis(
    name = "ae_forestly",
    label = "Interactive forest plot"
  ) |>
  meta_build()
```

```{r}
meta
```

After the metadata is defined, we can create an interactive forest plot for `wk12` as below.

```{r}
meta |>
  prepare_ae_forestly(
    population = "apat",
    observation = "wk12",
    parameter = "any;rel;ser"
  ) |>
  format_ae_forestly() |>
  ae_forestly()
```

Using similar approach, we could create an interactive forest plot for `period1`.

```{r}
meta |>
  prepare_ae_forestly(
    population = "apat",
    observation = "period1",
    parameter = "any;rel;ser"
  ) |>
  format_ae_forestly() |>
  ae_forestly()
```
