---
title: "Controlling Layout"
output:
  rmarkdown::html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
vignette: |
  %\VignetteIndexEntry{Controlling Layout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(forestly)
library(metalite)
```

# Overview

The forestly package controls forest plot layouts in a simple and effective approach.
To illustrate the idea, we use ADSL (`forestly_adsl_3grp`) and
ADAE (`forestly_adae_3grp`) datasets in the forestly package
that contain three treatment groups.

The first step is to create the required metadata using `meta_forestly()`.
Then we use `prepare_ae_forestly()` to prepare the input datasets for an interactive forest plot.
The `outdata` contains all the parameters required to generate the interactive forest plot.

```{r}
metadata <- meta_forestly(
  dataset_adsl = forestly_adsl_3grp,
  dataset_adae = forestly_adae_3grp,
  parameter_term = "any;rel;ser"
)

outdata <- metadata |> prepare_ae_forestly()

outdata
```

The interactive forest plot with default style can be generated.

```{r}
outdata |>
  format_ae_forestly() |>
  ae_forestly()
```

# Maximum Page Options 

By default, it will display the counts that rounding up to the nearest hundred displayed in the interactive forest table. 
It can be adjusted by adding "max_page =xx" argument to ae_forestly() function. 

````{r}
outdata |>
  format_ae_forestly() |>
  ae_forestly(max_page=280)

```


# Change color

By default, forestly is using teal for treatment group and plum for control group.
If user wants to change the color,
the `color` argument in the `format_ae_forestly()` function can be used.
Here is an example for an interactive forest plot using black and grey.

```{r}
outdata |>
  format_ae_forestly(color = c("black", "grey60", "grey40")) |>
  ae_forestly()
```

# Display different columns

By using the display argument in format_ae_forestly(), we can display the total column.

```{r}
outdata |>
  format_ae_forestly(display = c("n", "prop", "fig_prop", "fig_diff", "total")) |>
  ae_forestly()
```

We can also display risk difference columns in a similar approach.

```{r}
outdata |>
  format_ae_forestly(display = c("n", "prop", "fig_prop", "fig_diff", "diff")) |>
  ae_forestly()
```

# Control column width

We can control column width to customize the layout.

```{r}
outdata |>
  format_ae_forestly(
    width_fig = 230,
    footer_space = 110
  ) |>
  ae_forestly(width = 1000)
```

# Horizontal scrolling

Set `ae_forestly(..., width = NULL)` to make the element scrollable horizontally.
This is particularly useful when embedding the plot in pages with responsive
design, using frameworks such as Bootstrap.

# Change variable listed in drill-down table

Users can explore AE listing by clicking $\blacktriangleright$ of each row
after we specify column names in `ae_listing_display`.

```{r}
listing_var <- c(
  "SEX", "RACE", "AGE",
  "SITEID", "AESEV", "STUDYID",
  "AESER", "AEREL", "ASTDT", "AENDT"
)

metadata |>
  prepare_ae_forestly(
    ae_listing_display = listing_var
  ) |>
  format_ae_forestly() |>
  ae_forestly()
```

# Hide filters

You can use simple CSS rules to hide the incidence rate filter element.
For example, in an R Markdown document, use:

````{verbatim, lang="css"}
```{css, echo=FALSE}
#filter_subject {
  display: none;
}
```
````
