---
title: "A closer look at meta_forestly()"
output:
  rmarkdown::html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
vignette: |
  %\VignetteIndexEntry{A closer look at meta_forestly()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  echo = TRUE
)
```

```{r}
library(forestly)
library(metalite)
```

# Overview

`meta_forestly()` is the first step in the interactive forest plot generation workflow and is a wrapper of several functions from [metalite](https://merck.github.io/metalite/). Specifically, it defines a metadata, including ADaM datesets, population, observations and parameters to be used in the forest plot. 
In this vignette we will take a closer look at the `parameter_term` argument in `meta_forestly()` to better understand how to generate the forest plot for different parameters of interest.

# `parameter_term` examples

Prior to turning to the examples, just a reminder that the treatment group variable in the ADaM datasets should be set as a factor, otherwise a warning will be generated. 
For example, if this is what you have in the `TRT01A`:
```{r}
unique(forestly_adsl$TRT01A)
```

you can easily turn it into a factor variable with the desired labels using the following code: 

```{r}
forestly_adsl$TRTA <- factor(
  forestly_adsl$TRT01A, 
  levels = c("Xanomeline Low Dose", "Placebo"), 
  labels = c("Low Dose", "Placebo")
  )
```

A similar approach needs to be applied to the `TRTA`:

```{r}
forestly_adae$TRTA <- factor(
  forestly_adae$TRTA, 
  levels = c("Xanomeline Low Dose", "Placebo"), 
  labels = c("Low Dose", "Placebo")
  )
```

If the deliverable requires a forest plot of all adverse events and serious adverse events, then the following code could be used with `any` and `ser` parameters in `parameter_term` argument respectively:

```{r}
meta_forestly(
  dataset_adsl = forestly_adsl,
  dataset_adae = forestly_adae,
  parameter_term = "any;ser"
) |>
  prepare_ae_forestly() |>
  format_ae_forestly() |>
  ae_forestly()
```

However, if grade 3-5 adverse events need to be plotted as well, we can add `g35` parameter in the `parameter_term`. Click on the AE criteria below to see that a forest plot for grade 3-5 adverse events was added.

```{r}
meta_forestly(
  dataset_adsl = forestly_adsl,
  dataset_adae = forestly_adae,
  parameter_term = "any;ser;g35"
) |>
  prepare_ae_forestly() |>
  format_ae_forestly() |>
  ae_forestly()
```

# Adding a new parameter

Below we demonstrate how to add a new parameter to the `meta_forestly()`. Consider, for instance, that now we need to plot adverse events which resulted in dose reductions, a parameter that is currently not available in the [predefined options](https://merck.github.io/forestly/articles/forestly.html) of the `parameter_term`. Instead of using `meta_forestly()` directly, we will define `meta_new` object using [metalite](https://merck.github.io/metalite/) core functions as demonstrated below. 

Specifically, `define_parameter()` is where the new parameter `reduc` is defined by filtering `forestly_adae` with `AEACN == "DOSE REDUCED"` condition in order to identify the adverse events of interest:

```{r}
meta_new <- meta_adam(
  population = forestly_adsl,
  observation = forestly_adae
  ) |>
  define_plan(plan = metalite::plan(
    analysis = "ae_forestly",
    population = "apat",
    observation = "all",
    parameter = "any;ser;g35;reduc"
    )
  ) |>
  define_population(
    name = "apat",
    group = "TRTA",
  ) |>
  define_observation(
    name = "all",
    group = "TRTA",
    label = "All AEs"
  ) |>
  define_parameter(
    name = "reduc",
    subset = AEACN == "DOSE REDUCED",
    label = "adverse events resulting in dose reduction"
  ) |>
  define_analysis(
    name = "ae_forestly",
    label = "Interactive forest plot"
  ) |>
  meta_build()
```

Now, similar to the originally generated forest plot above, we need to apply `prepare_ae_forestly()`, `format_ae_forestly()`, and `ae_forestly()` to generate the plot that will include adverse events which resulted in dose reductions:

```{r}
meta_new |>
  prepare_ae_forestly() |>
  format_ae_forestly() |>
  ae_forestly()
```
